---
title: 'Gender or History of Anxiety Disorder Diagnosis: Which Factors May Contribute
  to Procrastination Patterns in People with Anxiety?'
author: "Nikita Nevskiy & Felisha Wade"
date: "`r Sys.Date()`"
output:
  html_document: default
---

# Part 0. First setup

```{r setup, include=FALSE}

# Loading the packages
library(pacman)
pacman::p_load(pacman, dplyr, GGally, ggplot2, ggthemes, ggvis, httr, lubridate, plotly, rio, rmarkdown, shiny, stringr, tidyr)
library(datasets)

# Basic codes for the first setup
rm(list = ls()) 
options(dplyr.width = Inf) 
options(scipen = 999) 

```

# Part I. Importing and filtering the data

```{r import_data, include=FALSE}

# Loading the data
APP <- read.csv("~/Downloads/APP.csv", header = T)

# Separating the results of two questionnaires
pswq_items <- APP[7:22] # Q1 to Q16
lay_items <- APP[23:42] # Q17 to Q36

```

# Part II. Reversal

```{r reversal, include=FALSE}

# Both questionnaires contain items to be reversed before performing the analyses
# Since both questionnaires use Likert-type scale (1-5), the reversed score is calculated using the formula "6 - x"
reverse_score <- function(data, item_numbers) {
  scored <- data
  scored[item_numbers] <- lapply(scored[item_numbers], function(x) 6 - x)
  scored
}

# PSWQ reverse-keyed items: 1, 3, 8, 10, and 11
pswq_items_rev <- reverse_score(pswq_items, c(1, 3, 8, 10, 11))

# Lay reverse-keyed items: 3, 4, 6, 8, 11, 13, 14, 15, 18, and 20 (or Q19, Q20, Q22, Q24, Q27, Q29, Q30, Q31, Q34, and Q36)
lay_items_rev <- reverse_score(lay_items, c(3, 4, 6, 8, 11, 13, 14, 15, 18, 20))

```

# Part III. Descriptive statistics

```{r descriptives, include=FALSE}

# First step is to calculate the mean values for both questionnaires
mean_values <- data.frame(
  participant_id = seq_len(nrow(APP)),
  age = APP[[3]],
  gender = APP[[4]],
  education = APP[[5]],
  anxiety_diagnosis = APP[[6]],
  anxiety_mean = rowMeans(pswq_items_rev),
  procrastination_mean = rowMeans(lay_items_rev)
)

# Setting the formula to compute Cronbach's alpha
cronbach_alpha <- function(items) {
  k <- ncol(items)
  k / (k - 1) * (1 - sum(vapply(items, var, numeric(1))) / var(rowSums(items)))
}

# The data set is now ready to present the descriptive statistics
descriptives <- data.frame(
  scale = c("PSWQ", "Lay's Procrastination Scale"),
  n = c(nrow(mean_values), nrow(mean_values)), # Sample size
  cronbach_alpha = c(
    cronbach_alpha(pswq_items_rev),
    cronbach_alpha(lay_items_rev)
  ),
  mean = c(mean(mean_values$anxiety_mean), mean(mean_values$procrastination_mean)),
  median = c(median(mean_values$anxiety_mean), median(mean_values$procrastination_mean)),
  sd = c(sd(mean_values$anxiety_mean), sd(mean_values$procrastination_mean)), # Standard deviation
  minimum = c(min(mean_values$anxiety_mean), min(mean_values$procrastination_mean)),
  maximum = c(max(mean_values$anxiety_mean), max(mean_values$procrastination_mean))
)

```

# Part IV. Pearson's correlation

```{r correlation, include=FALSE}

correlation_test <- cor.test(
  mean_values$anxiety_mean,
  mean_values$procrastination_mean,
  method = "pearson"
)

# The results of Person's correlation analysis, previously presented in form of the list, may now be demonstrated in a table.
correlation_result <- data.frame(
  variables = "results",
  r = unname(correlation_test$estimate), # Pearson's r
  df = unname(correlation_test$parameter), # Degrees of freedom
  p_value = correlation_test$p.value, # p-value
  n = nrow(mean_values)
)

```

# Part V. Visualising the results

## V.a. Table 1. Descriptive statistics for PSWQ and procrastination scales

```{r visuals_descriptives, echo=FALSE, message=FALSE}

knitr::kable(
  descriptives,
  digits = c(0, 0, 3, 2, 2, 3, 2, 2), # Cut the digits after comma
)

```

## V.b. Table 2. Correlation matrix for PSWQ and procrastination scales.

```{r visuals_cor, echo=FALSE, message=FALSE}

knitr::kable(
  correlation_result,
  digits = c(0, 3, 0, 3, 0),
)

```

## V.c. Figure 1. Scatter plot for correlation matrix

```{r visuals_sp_cor, echo=FALSE, message=FALSE}

ggplot(mean_values, aes(x = anxiety_mean, y = procrastination_mean)) +
  geom_point(size = 2, alpha = 0.75, colour = "#2C6E9F") +
  geom_smooth(method = "lm", se = TRUE, colour = "#C44E52", linewidth = 0.9) +
  labs(
    subtitle = sprintf(
      "Pearson's r = %.2f, p = %.3f",
      correlation_test$estimate,
      correlation_test$p.value
    ),
    x = "PSWQ",
    y = "Procrastination"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold"),
    panel.grid.minor = element_blank()
  )

```

# Part VI. Descriptive statistics by gender and history of anxiety diagnosis

```{r subgroup_descriptives, include=FALSE}

# Additional hypotheses (H1 and H2) imply separating the overall sample first by gender, and second — by history of anxiety diagnosis

# Gender is available for all 56 participants
gender_descriptives <- mean_values %>%
  group_by(gender) %>%
  summarise(
    n = n(),
    PSWQ_M = mean(anxiety_mean),
    PSWQ_Med = median(anxiety_mean),
    PSWQ_SD = sd(anxiety_mean),
    PSWQ_min = min(anxiety_mean),
    PSWQ_max = max(anxiety_mean),
    Proc_M = mean(procrastination_mean),
    Proc_Med = median(procrastination_mean),
    Proc_SD = sd(procrastination_mean),
    Proc_min = min(procrastination_mean),
    Proc_max = max(procrastination_mean),
    .groups = "drop"
  )

# One participant selected "Prefer not to say" for diagnostic history. This response cannot be assigned to either comparison group, so it is excluded from this subgroup analysis.
diagnosis_descriptives <- mean_values %>%
  filter(anxiety_diagnosis %in% c("Yes", "No")) %>%
  group_by(anxiety_diagnosis) %>%
  summarise(
    n = n(),
    PSWQ_M = mean(anxiety_mean),
    PSWQ_Med = median(anxiety_mean),
    PSWQ_SD = sd(anxiety_mean),
    PSWQ_min = min(anxiety_mean),
    PSWQ_max = max(anxiety_mean),
    Proc_M = mean(procrastination_mean),
    Proc_Med = median(procrastination_mean),
    Proc_SD = sd(procrastination_mean),
    Proc_min = min(procrastination_mean),
    Proc_max = max(procrastination_mean),
    .groups = "drop"
  )

diagnosis_data <- mean_values %>%
  filter(anxiety_diagnosis %in% c("Yes", "No"))

```

# Part VII. Visualising the separated results

## VII.a. Table 3. Descriptive statistics divided by gender.

```{r gender_table, echo=FALSE}

knitr::kable(
  gender_descriptives,
  digits = c(0, 0, 2, 2, 3, 2, 2, 2, 2, 3, 2, 2),
)

```

## VII.b. Figure 2. Scatter plot for descriptive statistics divided by gender

```{r gender_scatter, echo=FALSE, message=FALSE, fig.width=7, fig.height=5}

ggplot(mean_values, aes(x = anxiety_mean, y = procrastination_mean)) +
  geom_point(aes(colour = gender), size = 2, alpha = 0.8) +
  geom_smooth(method = "lm", se = TRUE, colour = "#555555", linewidth = 0.9) +
  labs(
    subtitle = "The regression line represents the full sample; the shading around it is the standard error.",
    x = "PSWQ",
    y = "Procrastination",
    colour = "Gender"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold"),
    panel.grid.minor = element_blank()
  )

```

## VII.c. Table 4. Descriptive statistics divided by history of anxiety diagnosis.

```{r diagnosis_table, echo=FALSE}

knitr::kable(
  diagnosis_descriptives,
  digits = c(0, 0, 2, 2, 3, 2, 2, 2, 2, 3, 2, 2),
  caption = "One N/A response excluded."
)

```

## VII.d. Figure 3. Scatter plot for descriptive statistics divided by history of anxiety diagnosis.

```{r diagnosis_scatter, echo=FALSE, message=FALSE, fig.width=7, fig.height=5}

ggplot(diagnosis_data, aes(x = anxiety_mean, y = procrastination_mean)) +
  geom_point(aes(colour = anxiety_diagnosis), size = 2.3, alpha = 0.8) +
  geom_smooth(method = "lm", se = TRUE, colour = "#555555", linewidth = 0.9) +
  labs(
    subtitle = "One N/A response is excluded. The regression line represents the full sample; the shading around it is the standard error.",
    x = "PSWQ",
    y = "Procrastination",
    colour = "Diagnosis history"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold"),
    panel.grid.minor = element_blank()
  )

```
